---
title: "CIMseq Plotting Vignette"
author: "Jason T. Serviss, Nathanael Andrews, and Martin Enge"
date: '`r Sys.Date()`'
output:
    html_document:
        theme: flatly
        toc: yes
vignette: >
    %\VignetteIndexEntry{CIMseq Plotting Vignette} 
    %\usepackage[utf8]{inputenc}
    %\SweaveUTF8
    %\VignetteEngine{knitr::rmarkdown} 
editor_options: 
  chunk_output_type: console
---

```{r, message = FALSE}
packages <- c(
  "CIMseq", "printr", "ggthemes", "dplyr", 
  "tidyr", "ggplot2", "viridis"
)
purrr::walk(packages, library, character.only = TRUE)
rm(packages)
```

## CIMseqSinglets and CIMseqMultiplets

Plot markers.
```{r, fig.align="center", fig.width=10, fig.height=8, eval = FALSE}
plotCountsMarkers(
  CIMseqSinglets_test, CIMseqMultiplets_test, markers = c("CD74", "ANXA3")
)
```

Plot cell number.
```{r, fig.align="center", fig.width=10, fig.height=8}
plotCountsERCC(CIMseqSinglets_test, CIMseqMultiplets_test)
```

Get the plotting data so you can make a custom plot.
```{r, fig.align="center", fig.width=10, fig.height=8}
plotCountsMarkers(
  CIMseqSinglets_test, CIMseqMultiplets_test, markers = c("CD74", "ANXA3")
) %>%
  plotData() %>%
  head()

plotCountsERCC(CIMseqSinglets_test, CIMseqMultiplets_test) %>%
  plotData() %>%
  head()
```

Plot the clustering and classification.
```{r, fig.align="center", fig.width=10, fig.height=8}
plotUnsupervisedClass(CIMseqSinglets_test, CIMseqMultiplets_test)
```

Plot a single marker.
```{r, fig.align="center", fig.width=10, fig.height=8}
plotUnsupervisedMarkers(CIMseqSinglets_test, CIMseqMultiplets_test, "CD74")
```

Plot markers in different cell types.
```{r, fig.align="center", fig.width=10, fig.height=8}
plotUnsupervisedMarkers(
  CIMseqSinglets_test, CIMseqMultiplets_test,
  markers = c("CD74", "ANXA3", "ACTG2")
)
```

Plot mean of multiple markers for the same cell type.
```{r, fig.align="center", fig.width=10, fig.height=8, message = FALSE}
#use the data instead
plotUnsupervisedMarkers(
  CIMseqSinglets_test, CIMseqMultiplets_test, 
  markers = c("CD74", "HLA-DRA", "IL13RA2", "MAGEA4")
) %>%
  plotData() %>%
  gather(gene, value, -Sample, -(`Sample type`:Colour)) %>%
  group_by(Sample) %>%
  mutate(`Mean(markers)` = mean(value)) %>%
  ungroup() %>%
  select(`dim.red dim 1`, `dim.red dim 2`, `Mean(markers)`) %>%
  distinct() %>%
  ggplot() +
  geom_point(aes(`dim.red dim 1`, `dim.red dim 2`, colour = `Mean(markers)`)) +
  viridis::scale_colour_viridis(option = "E") +
  theme_few() +
  theme(legend.position = "top", legend.title.align = 0) +
  guides(colour = guide_colourbar(title.position = "top"))
```

Get the plotting data so you can make a custom plot.

```{r}
plotUnsupervisedClass(CIMseqSinglets_test, CIMseqMultiplets_test) %>% 
  plotData() %>%
  head()

plotUnsupervisedMarkers(
  CIMseqSinglets_test, CIMseqMultiplets_test, 
  markers = c("CD74", "ANXA3", "ACTG2")
) %>% 
  plotData %>%
  head()
```

## CIMseqSwarm

### Graph-based plots

```{r, fig.align="center", fig.width=10, fig.height=8}
plotSwarmGraph(CIMseqSwarm_test, CIMseqSinglets_test, CIMseqMultiplets_test)
```

### Non-graph-based plots

Edge bar graph.
```{r, fig.align="center", fig.width=10, fig.height=8}
plotSwarmEdgeBar(CIMseqSwarm_test, CIMseqSinglets_test, CIMseqMultiplets_test)
```

P-value bar graph.
```{r, fig.align="center", fig.width=10, fig.height=8}
plotSwarmPbar(CIMseqSwarm_test, CIMseqSinglets_test, CIMseqMultiplets_test)
```

Heatmap.
```{r, fig.align="center", fig.width=10, fig.height=8}
plotSwarmHeat(CIMseqSwarm_test, CIMseqSinglets_test, CIMseqMultiplets_test)
```

```{r, fig.align="center", fig.width=10, fig.height=8}
plotSwarmCircos(
  CIMseqSwarm_test, CIMseqSinglets_test, CIMseqMultiplets_test, alpha = 1
)
```

Get data to make a custom plot.
```{r}
# plotSwarmGraph(CIMseqSwarm_test, CIMseqSinglets_test, CIMseqMultiplets_test) %>%
#  plotData()

plotSwarmEdgeBar(CIMseqSwarm_test, CIMseqSinglets_test, CIMseqMultiplets_test) %>%
  plotData() %>%
  head()
```


### Gene plots
```{r, fig.align="center", fig.width=10, fig.height=8, eval = FALSE}
#taking too long... why?
plotSwarmGenes(
  CIMseqSwarm_test, CIMseqSinglets_test, CIMseqMultiplets_test,
  c("CD74", "ANXA3", "ACTG2"), rownames(getData(CIMseqSwarm_test, "fractions"))
)
```

